Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-07 paper_jasa.tex
lots of cleanup everywhere, division of results section

However, measuring the strength of patterns seen in plots, and differences in individual perceptual ability, is something that is difficult and perhaps handicaps graphics  use among statisticians, where measuring probabilities is of primary importance. This has also been addressed in recent research. \citet{buja:2009} proposes a protocol that allow the testing of discoveries made from statistical graphics. This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis. One of the protocols, the lineup, places the data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different. Figure \ref{fig:test_category} shows an example lineup. Which plot do you think is the most different from the others? (The position of the actual data plot is provided in Section \ref{sec:category}.) Wrapped in a process that mirrors classical inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis. The null hypothesis typically derives from the task at hand, or the type of plot being made. The alternative encompasses all possible antitheses, all types of patterns that might be detected in the data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time. The probability of rejection can be quantified, along with Type I, and Type II error, and $p$-value and power can be defined and estimated.

@@

Table \ref{tbl:compare} illustrates the lineup protocol in relation to traditional hypothesis testing. Both start from the same place, the same null hypothesis. The classical test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error. In the lineup protocol, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values. In traditional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true. If it is extreme on this scale then the null hypothesis is rejected. In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution. If the data plot is selected as the most different, then this results in rejection of the null hypothesis.

@@

\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate observed data. \\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for p-value.  \\ Should I change everything accordingly? }

@@

We need to make two crucial assumptions:
\begin{enumerate}
\item  the plot setup is structured in a way that makes it possible for observer to identify a deviation from the null hypothesis,
\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
\end{enumerate}
We will be able to show evidence for the second assumption with the help of the study discussed in section~\ref{sec:simulation}, \hh{the degree to which the first assumption is fulfilled is reflected} by the power of a lineup. The better suited a design is for a particular task, the higher its power will be.

@@

Figure \ref{fig:pval_power} gives an overview of the probability of picking the data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance. We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the data plot from null plots in a lineup of size $m=20$.  This pattern is confirmed by our findings described in section~ref{sec:simulation}
%also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3.5in]{images/powerplot.pdf} 
   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_D$. The grey line shows probability of picking the data plot under $H_0$. }
   \label{fig:pval_power}
\end{figure}

Now assume that the same lineup plot is evaluated by $K$ independent observers. W.l.o.g. let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.  Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$. We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.

@@

Note that the right hand side of the equation (\ref{eqn:power_estimate}) is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers. While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.

@@

The classical test in this scenario uses $T= \hat{\beta_k}/
se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme
on the scale of a $t$ distribution with $n-p$ degrees of freedom. This
is the uniformly most powerful (UMP) test \green{reference needed
  here} and it forms the benchmark upon which we evaluate the visual
test. To calculate what we might expect for the power of the visual
test, under perfect conditions, first assume that the observer is able
to pick the plot with the smallest $p$-value from a lineup plot.  This
leads to the decision to reject $H_0$ when $p_{B} < p_0$, where $p_{B}$ is calculated according to details given in Lemma \ref{lemma}. Thus 
the expected power of the visual test in this scenario is

@@

\section{Simulation Experiments} \label{sec:simulation}

@@

@@

\green{Discuss the UMP test \blue{(discussed right after equation \ref{power_exp}) }, and give the reference for it here \blue { I am looking for a good reference. Do you have any suggestion?}.}

@@

This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2. Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$. For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups. \green{ I don't understand the next two sentences.} For some parameter combinations we have more than three replications \blue{(since we run an extension of this experiment with more lineups for some parameter combinations}. Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots \blue{(we can ignore this now as I see that blocking discussion in section \ref{sec:category} is removed.) }.

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11. \green{Check this} \blue{(Yes, it is 11 !)} This lineup, we think, is a difficult one. To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest). For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized.\blue{The order is as described not randomized after. The randomization is done picking first combination of sample size and sd as well as picking lineups for each combination of parameter} \green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups. The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants. It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. \green{perhaps should go into so much detail about this, or keep this for the conclusions section.}

@@

\green{Discuss the UMP test, and give the reference for it here.}

@@

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data. For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points. Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen such that overall error standard deviation of the data is $5$ which is approximately 3.5\green{????} \blue{(Yes, this is how we did it. This value (3.5) makes the overall sd to be 5 because of the inclusion of the contaminated part of the data.)}. Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.

@@

An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? Plot 7 is the actual data plot.

@@

\begin{figure}[hbt]
%\begin{figurehere}
   \centering
       \scalebox{0.4}{\includegraphics{contaminated_data.pdf}}
       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
}
       \label{fig:cont_dat}
\end{figure}

\subsection{Results}

Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk. Summary of the data we obtained for all three experiments are shown in table \ref{tbl:summary}. Participants are asked to select the plot they think best match the question asked, give a reason for their choice, and determine a level of  confidence for their decision. Gender, age, education and geographic location of each participant are also collected. In this process of collecting data some participants did not provide demographic information as we see in table \ref{tbl:summary} that male and female participants do not add up to total participants. It is not clear why they did not provide the information. Now the concern arises whether we should keep their responses in the study. It is also possible that participants were not serious enough to provide feedback. For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.  This could be a criteria to determine their attentiveness in the study. For this we consider five different data cleaning criteria as shown below;

@@

\begin{table}[hbtp]
\caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
\begin{center}
\begin{tabular}{rrrrr|rrrr|rrrr}
  \hline
Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
 \cline{2-5}  \cline{6-9}   \cline{10-13}
criteria & total & male & female & feedbacks & total & male & female & feedbacks & total & male & female & feedbacks \\ 
  \hline
1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
       \caption{The subject wise power estimated from model \ref{mixed} are shown for various data sets obtained by applying 5 screening criteria.  The corresponding power curve for Uniformly Most powerful (UMP) test is shown in pink lines for comparison.}
       \label{fig:power_screening_subject}
\end{figure}

Each participant was shown a sequence of 10 lineup plots.  In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.  The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using equation \eqref{power_est} along with 95\% confidence interval calculated using Fisher's exact method.

@@

%We fit model \eqref{mixed} to the survey data obtained from the simulation experiment. The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}. Model \ref{mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}. The plot includes 20 randomly selected subject-specific power curves. Notice that the power curve estimated for one subject is above the UMP test power curve. 
%
%\begin{figure}[hbtp]
%   \centering
%       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
%       \caption{Estimated subject specific power curve from model \ref{mixed} for sample size $n$ = 100 and $\sigma$ = 12.  The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
%       \label{fig:power_subject}
%\end{figure}

@@

\hh{From figure \ref{fig:power_loess3} we see that identification of the data plot is high independently of the effect size. This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e. the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of `correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
       \label{fig:power_observed}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
       \label{fig:power_loess1}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
       \label{fig:power_loess2}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
       \caption{Subject specific power obtained from equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
       \label{fig:power_effect_subject1}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in table \ref{tbl:experiment_params}.}
       \label{fig:power_diff_exp}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
       \caption{Observed power of visual test from equation \ref{power_est} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3. }
       \label{fig:power_observed_exp3}
\end{figure*}

\green{A lot more needs to be done here!!!}

>>>>

However, measuring the strength of patterns seen in plots, and differences in individual perceptual ability, is something that is difficult and perhaps handicaps graphics  use among statisticians, where measuring probabilities is of primary importance. This has also been addressed in recent research. \citet{buja:2009} proposes a protocol that allow the testing of discoveries made from statistical graphics. This work represents a major advance for graphics, because it bridges the gulf between classical statistical inference procedures and exploratory data analysis. One of the protocols, the lineup, places the actual data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different. Figure \ref{fig:test_category} shows an example lineup. Which plot do you think is the most different from the others? (The position of the actual data plot is provided in Section \ref{sec:category}.) Wrapped in a process that mirrors classical inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis. The null hypothesis typically derives from the task at hand, or the type of plot being made. The alternative encompasses all possible antitheses, all types of patterns that might be detected in the actual data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time. The probability of rejection can be quantified, along with Type I, and Type II error, and $p$-value and power can be defined and estimated.

@@

Table \ref{tbl:compare} illustrates the lineup protocol in relation to traditional hypothesis testing. Both start from the same place, the same null hypothesis. The classical test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error. In the lineup protocol, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values. In traditional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true. If it is extreme on this scale then the null hypothesis is rejected. In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution. If the actual data plot is selected as the most different, then this results in rejection of the null hypothesis.

@@

\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate observed data. \\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for p-value.  \\ Should I change everything accordingly? I see that we have been careful not to use $\beta$ for power! }

@@

We need to make two crucial assumptions:
\begin{enumerate}
\item  the plot setup is structured in a way that makes it possible for observer to identify a deviation from the null hypothesis,
\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
\end{enumerate}
We will be able to show evidence for the second assumption with the help of the study discussed in Section~\ref{sec:simulation}, \hh{the degree to which the first assumption is fulfilled is reflected} by the power of a lineup. The better suited a design is for a particular task, the higher its power will be.

@@

Figure \ref{fig:pval_power} gives an overview of the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance. We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the actual data plot from null plots in a lineup of size $m=20$.  This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
%also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3.8in]{images/powerplot.pdf} 
   \caption{Comparison of probabilities to pick the actual data plot for different size lineups, $m$, and different values for data plot strength, $p_D$. The dashed  line, and grey 95\% pointwise confidence band, shows the probability of picking the data plot, by chance, if $H_0$ is true. }
   \label{fig:pval_power}
\end{figure}

Now assume that the same lineup plot is evaluated by $K$ independent observers. W.l.o.g.\green{Can we get rid of cliches like WLOG?} let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.  Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$. We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.

@@

Note that the right hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers. While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.

@@

The classical test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom. This is the uniformly most powerful (UMP) test \green{reference needed here} and it forms the benchmark upon which we evaluate the visual test. (This is also the UMP test for the continuous covariate, case 2.) To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.  This leads to the decision to reject $H_0$ when $p_{B} < p_0$, where $p_{B}$ is calculated according to details given in Lemma \ref{lemma}. Thus the expected power of the visual test in this scenario is

@@

\section{Human Subjects Experiments with Simulated Data} \label{sec:simulation}

@@

Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk. A summary of the data obtained for all three experiments are shown in Table \ref{tbl:summary}. Participants were asked to select the plot they think best matched the question given, provide a reason for their choice, and say how confident they are in their choice. Gender, age, education and geographic location of each participant are also collected. The data was collected through the web. For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.

@@

@@

This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2. Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$. For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups. For some parameter combinations we have more than three replications, because additional lineups were produced for values of the parameters that exhibited large variation in the initial data.

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11. This lineup, we think, is a difficult one. To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest). For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized. Each subject saw a substantial number of difficult lineups. The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants. It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. \green{perhaps should go into so much detail about this, or keep this for the conclusions section.}

@@

@@

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data. For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points. Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$. Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.

@@

An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? %Plot 7 is the actual data plot. 
\begin{figure}[hbt]
%\begin{figurehere}
   \centering
       \scalebox{0.4}{\includegraphics{contaminated_data.pdf}}
       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
}
       \label{fig:cont_dat}
\end{figure}

@@

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
%       \caption{Observed power of visual test from Equation \ref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
%       \label{fig:power_observed}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
%       \label{fig:power_loess1}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
%       \label{fig:power_loess2}
%\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
       \caption{Subject specific power obtained from Equation \ref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
       \label{fig:power_effect_subject1}
\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
%       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in table \ref{tbl:experiment_params}.}
%       \label{fig:power_diff_exp}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
%       \caption{Observed power of visual test from Equation \ref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3. }
%       \label{fig:power_observed_exp3}
%\end{figure*}

\section{Results}

\subsection{Data Cleaning}

% In this process of collecting data some participants did not provide demographic information as we see in Table \ref{tbl:summary} that male and female participants do not add up to total participants. It is not clear why they did not provide the information. Now the concern arises whether we should keep their responses in the study. It is also possible that participants were not serious enough to provide feedback. 
For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.  This could be a criteria to determine their attentiveness in the study. For this we consider five different data cleaning criteria as shown below;

@@

\begin{table}[hbtp]
\caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
\begin{center}
\begin{tabular}{rrrrr|rrrr|rrrr}
  \hline
Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
 \cline{2-5}  \cline{6-9}   \cline{10-13}
criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
  \hline
1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

\subsection{Power comparison}

@@

Each participant was shown a sequence of 10 lineup plots.  In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.  The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using Equation \ref{eqn:power_estimate} along with 95\% confidence interval calculated using Fisher's exact method.

@@

%We fit model \eqref{mixed} to the survey data obtained from the simulation experiment. The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}. Model \ref{mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}. The plot includes 20 randomly selected subject-specific power curves. Notice that the power curve estimated for one subject is above the UMP test power curve.

%\begin{figure}[hbtp]
%   \centering
%       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
%       \caption{Estimated subject specific power curve from model \ref{mixed} for sample size $n$ = 100 and $\sigma$ = 12.  The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
%       \label{fig:power_subject}
%\end{figure}

\hh{From figure \ref{fig:power_loess3} we see that identification of the data plot is high independently of the effect size. This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e. the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of `correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}.} \green{Not sure that this goes here, but something needs to be said about Type III.}

\subsection{Subject-specific variation}

\subsection{Estimating $p$-value in on-Classical Settings}

@@
